\section{Document History}
	\subsection{Pending Changes}
		As of 8/27/2022 the following changes are waiting to be approved by the Board of Trustees:
		\begin{itemize}
			\item{\change{Updates to Faculty Handbook from Personnel Committee}{30}{}}
			\item{\change{Sexual Harassment etc. policies dictated by CA law}{5}{}}
			\item{\change{Identify founder with her name: Waiting for BoT approval}{24}{}}
		\end{itemize}

		As of 8/27/2022 the following changes are waiting to be approved by the Provost/Faculty Council:
		\begin{itemize}
			\item{\change{Give paragraph numbers to titles that are lacking them}{32}{}}
			\item{\change{Fix the numbering and formatting errors associated with the September 2020 version of the Handbook}{23}{}}
			\item{\change{Org chart referred to in 1.3.2 doesn't exist}{26}{}}
		\end{itemize}


	\subsection{History}
		\begin{itemize}
			\item{{\bfseries 8/2022} Incorporated the pending changes from \change{issue 30 approved at faculty meeting in spring 2022}{30}{}}
			\item{{\bfseries 8/2022} An analysis of the differences between the 8/2021 and 1/2022 version was conducted.
				The changes were to the sexual harassment policy in
				\ref{sec:PolicyOnHarassment}
				The text was incorporated into this document on 8/27/2022.
				\change{Changes}{28}{Text of changes was taken from the 1/2022 wild copy}
				were discussed with Faculty Council in 2021-2022 school by VP of student life, E. Schultze and incorporated by virtue of 1.5.3.2.1
			}
			\item{{\bfseries 8/2022} An analysis of the differences between the 9/2020 and 1/2022 version was conducted. No substantial differences were noted.}
			\item{{\bfseries 6/2022} Untracked version of the Handbook were identified in the wild dated 1/2022}
			\item{{\bfseries 3/2022} The transcription from the September 24,2020 was completed.}
			\item{{\bfseries 12/2021} The transcription from the September 24,2020 was completed.}
			\item{{\bfseries 3/2021} Untracked version of the Handbook were identified in the wild dated 8/2021}
			\item{{\bfseries 12/2020} Transcription was begun from a Word Document, dated September 24,2020.}
		\end{itemize}