

\section{ACADEMIC SERVICES}
	\subsection{Library \& Information Services}
		Roger John Voskuyl Library, named for Westmont's third president, provides access to information resources and services to support the research and information needs of faculty, staff, students and the surrounding community.  The library collections include books and e-books, media items, music scores, microforms, print periodicals, and online resources including periodicals, databases and other online tools.  The Westmont College Archives collection is located on the lower level of the library and is comprised of items relating to its history, 1940 to present, and that of its predecessors, Western Bible College, 1939-1940, and the Bible Missionary Institute, 1937-1939.
		\begin{enumerate}[label=\alph*)]
			\item{\underline{Library Facility}:  The Learning Commons is located on the main level of Voskuyl Library.  This space brings together library, technology and other campus services in an environment designed to foster collaborative and creative work, and social interaction.  The learning spaces in the library are designed for different types of learning styles, including quiet area on the upper and lower levels, conference rooms, an open computer center, a technology teaching lab, group study rooms, individual study stations, group study tables and comfortable lounge furniture.
			}
			\item{\underline{Services}:  Library Services available for faculty include interlibrary loan, course reserves, the department liaison program, and the information literacy instruction program.  Detailed information about the library's services and programs may be found on the library website: 
		\href{
			http://library.westmont.edu
		}{\url{
				http://library.westmont.edu
			}}
			\item{\underline{Library Instruction}:  Guidance in the use of the library resources is provided by subject and course-related research guides available on the library website.  Information literacy instruction sessions, taught by Westmont librarians, are offered in the library technology teaching lab.  Librarians are also available for in-class instruction sessions as well as one-on-one research consultations.
			}
			\item{\underline{Loans}:  Faculty may borrow materials from the Voskuyl's collections for a period determined by and subject to the library's circulation policy.  Faculty have access to additional resources through interlibrary loan.
			}
			\item{\underline{Collection Development}:  Collection development is seen as a cooperative effort between library professionals and teaching faculty in various academic departments.
			}
			\item{\underline{Cubicles}:  A limited number of library cubicles are available to faculty who are involved in special projects.  Applications for cubicles are made through the Professional Development Committee.
			}
		\end{enumerate}/policy/

	\subsection{Computer Services}

		\underline{Information Technology Computing and Network Services}:  Information technology provides a wide variety of computing and network services in support of the college's mission.  These services include:  local network infrastructure; connection to the Internet; servers for email, web pages, file serving and applications; workstations for faculty and staff offices; data projection systems and workstations in classrooms; campus-wide licensing for general-use software; a course management system; training and assistance in developing resources such as the ones mentioned above.

		Campus network users are expected to abide by our Resource Use Policy found at
		\href{
			https://www.westmont.edu/information-technology/resource-use-policy
		}{\url{
				https://www.westmont.edu/information-technology/resource-use-policy
			}}
		.

		Problems involving any service provided by IT may be reported via the web at
		\href{
			https://support.westmont.edu/
		}{\url{
				https://support.westmont.edu/
			}}
		,
		or via email at support@westmont.edu
		,
		or by phone at 565-7211.  Further information about these services may be found at
		\href{
			https://www.westmont.edu/information-technology
		}{\url{
				https://www.westmont.edu/information-technology
			}}
		.
